---
title: "apsimx: An R package for APSIM Next Generation"
author: "Fernando Miguez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{apsimx: An R package for APSIM Next Generation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(apsimx)
library(ggplot2)
```

# Introduction

Are you a user of APSIM-X ("Next Generation")? Are you a frequent user of R? Are you hoping for APSIM-X and R to interact with each other? Then the answer is the **apsimx** package, which allows you to inspect, edit, run and read files created by APSIM-X (.apsimx) files. New .apsimx files are based on the JSON (JavaScript Object Notation https://www.json.org/) format and it is used to communicate information to the APSIM-X engine. By using the 'jsonlite' R package **apsimx** can read and write ('inspect' and 'edit') files in this format. The older format based on XML is partially supported through the 'xml2' package. To run simulations the *apsimx* function is available and results from the simulation can be imported as an R object with a dedicated function (*read_apsimx*). If you need to write scripts (regardless of whether you use the **apsimx** package or not) the vignette 'writing scripts' might be useful. The package is currently being developed and available at: https://github.com/femiguez/apsimx. 
If you have any questions write to: femiguez *at* iastate.edu. 

### Additional information

Before using this package it is recommended to become familiar with APSIM-X ("Next Generation").

See the following for details

[APSIM](https://www.apsim.info)

[APSIM-X](https://apsimnextgeneration.netlify.com/)

[Holzworth et al. (2018)](https://www.sciencedirect.com/science/article/pii/S1364815217311921)

This package has been tested on:

**Mac**: ApsimX version 4035

**Debian**: ApsimX version 4000

**Windows**: ApsimX version 4025

Other useful R packages are 'APSIM' and 'apsimr'. The main difference is that these were written to work with the previous generation of APSIM (7.x). This package is thus an update for the new APSIM-X software.

## Installing APSIM-X

For most functions in this pacakge you need to have a version of APSIM-X installed. For this you might first need to install mono (https://www.mono-project.com/download/stable/) and then download and install APSIM-X (https://www.apsim.info/Products/Downloads.aspx). The \code{apsimx} package will try to detect the install of both mono and APSIM-X. 

**Mac**: For Mac it is also assumed that APSIM-X has been moved to 'Applications'.

**Windows**: In Windows the default install location is in 'Program Files'.

## R functions for APSIM-X

The main functions in the package are (for now):

* **inspect_apsimx** simple inspecting of an .apsimx file. (It does not replace the GUI.)
* **edit_apsimx** edit/modify an .apsimx file. (Limited options are available at the moment).
* **apsimx** run an .apsimx simulation.
* **read_apsimx** read files created by an .apsimx simulation run (.db files)
* **read_apsimx_all** read all files created by one or more .apsimx simulation runs in a directory (.db files)
* **inspect_apsimx_replacement** inspect a replacement component of an .apsimx file (JSON only)

### 'apsimx_exmple': Running a built-in example

To get started you can run one of the examples distributed with APSIM-X. This function does the following: it detects where the APSIM examples are located, it runs the example indicated and it returns the 'report' as a data.frame. After that you can do the usual manipulations, including visualization.

```{r apsimx-example, eval = TRUE}
maize <- apsimx_example("Maize")
summary(maize)
## Simple data plotting
ggplot(data = maize , aes(x = Date, y = Maize.AboveGround.Wt)) + 
  geom_point()
```

### 'inspect_apsimx': Inspecting a file

A function to inspect an .apsimx file is included. This is meant to be used simply to avoid the need to use the GUI when a quick inspection of a given component is needed. For example, if you edit a file and want to make sure that the intended component was edited correctly just inspect it. For complex inspections and manipulations of an .apsimx file the GUI is recommended.

```{r inspect-apsimx}
ex.dir <- auto_detect_apsimx_examples()
inspect_apsimx("Maize", src.dir = ex.dir, node = "Weather")
```

### 'edit_apsimx': Editing a file

Editing a file has the side effect of creating a new file with '-edited' added to the name to avoid conflict or unintended consequences, but it is also possible to overwrite it. This follows the same style as in the 'apsimr' package. This code does not run an APSIM simulation, it only edits the file. 

Let's say we want to modify the values of 'Organic Carbon' in the soil in the 'Maize.apsimx' example. First we should inspect the current values.

```{r inspect-maize-oc}
inspect_apsimx("Maize.apsimx", src.dir = ex.dir, node = "Soil",
               soil.child = "OrganicMatter")
```

The function 'inspect_apsimx', in this example, displays various information about the organic matter in the soil. (See the help page for more options). In the next step we 'edit' or change the values.

```{r edit-maize-oc}
ocs <- c(1.5,1.4,1.3,1.2,1.1,1.0,0.9)
edit_apsimx("Maize.apsimx", src.dir = ex.dir,
             wrt.dir = ".",
             node = "Soil",
             soil.child = "OrganicMatter", 
             parm = "OC", value = ocs)
```

Let's make sure that we have changed the values successfully.

```{r inspect-maize-oc-check}
inspect_apsimx("Maize-edited.apsimx", src.dir = ".", 
               node = "Soil",
               soil.child = "OrganicMatter")
```

Deleting the file.

```{r edit-apsimx-delete, echo = TRUE}
## delete the created file
file.remove("./Maize-edited.apsimx")
```

## Running a simulation

The proposed workflow is to start by creating an .apsimx file using the GUI. In following steps the goal could be, for example, to either perform sensitivity or uncertainty analysis using R. The funcitons included in this package are aimed at simplifying the process of running a simulation, changing the value of a variable/parameter and running the model again iteratively. Writing complex scripts requires a high level knowledge of both APSIM-X and R, but it is made possible with these tools. This is the proposed workflow:

1. Create a simualtion using the APSIM GUI
2. Identify variables/parameters that you want to modify (they need to be exposed in the .apsimx XML o JSON file)
3. Edit the file using 'edit_apsimx'
4. Run the simulation using 'apsimx'
5. Repeat steps 3 and 4 as needed
6. Read in the results, possibly using 'read_apsimx'
7. Perform visual and statistical analysis on the results

Another proposed more advanced workflow would involve calibrating or optimizing variables given observed data. Additional tools for implementing this type of work will be added in future development.

```{r apsimx}
## One example ('Wheat') is included with the package for the vignette
ex.dir <- system.file("extdata", package = "apsimx")
sim <- apsimx("Wheat.apsimx", src.dir = ex.dir, value = "report")
## Calcualte summary statistics on all variables
summary(sim)
## Plot data
ggplot(data = sim, aes(x = Date, y = Yield)) + geom_point()
## Inspect the Wheat .apsimx file
inspect_apsimx("Wheat", src.dir = ex.dir, node = "Crop")
inspect_apsimx("Wheat", src.dir = ex.dir, node = "Manager")
```

## Inspecting a 'Replacement' component

Many additional options in an APSIM-X simulation can be realized through the use of 'Replacements'. For this, there is a function that allows inspection. For this, it is important to know the structure of the 'replacement' but if exploring them, the available components will be displayed. For example, if we want to display the xy pair for the thermal time in the phenology component of the 'Maize' replacement we might go through the following steps:

State the node ("Maize") and display available components:

```{r inspect-replacement-node}
extd.dir <- system.file("extdata", package = "apsimx")
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                            node = "Maize", display.available = TRUE)
```

Choose 'Phenology' as a 'child' of 'Maize':

```{r inspect-replacement-node-child}
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                            node = "Maize", node.child = "Phenology",
                           display.available = TRUE)
```

Choose 'ThermalTime' as a 'sub.child' of 'Phenology':

```{r inspect-replacement-node-subchild}
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                            node = "Maize", node.child = "Phenology",
                           node.subchild = "ThermalTime",
                           display.available = TRUE)
```

Choose the 'X' and 'Y' parameters in the 'ThermalTime' sub.child.

```{r inspect-replacement-node-subchild-parm}
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                            node = "Maize", node.child = "Phenology",
                            node.subchild = "ThermalTime", parm = c("X","Y")) 
```

The function shows the available replacements ("Maize" and "Soybean"), the 'CropType' if available, the subchild name and the xy pair in this example. If parm is not specified all elements will be displayed.

Let's say we want to inspect details of soybean cultivars in the same file:

```{r inspect-replacement-soybean-cultivar-node}
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                            node = "Soybean", display.available = TRUE) 
```

In this case the cultivars are at the sub-node level.

```{r inspect-replacement-soybean-cultivar-node-child}
inspect_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                           node = "Soybean", 
                           node.child = "PioneerP22T61_MG22",
                           display.available = FALSE) 
```

## Editing a 'Replacement' 

After using the 'inspect' version of the file editing should be much more straight forward.

```{r edit-replacement}
edit_apsimx_replacement("MaizeSoybean.apsimx", src.dir = extd.dir,
                        node = "Maize", node.child = "Phenology",
                        node.subchild = "ThermalTime", node.subsubchild = "BaseThermalTime",
                        parm = "Y", value = c(0,12,20,28,0))
```

Let's inspect the edited file

```{r inspect-edit-replacement}
inspect_apsimx_replacement("MaizeSoybean-edited.apsimx", src.dir = ".",
                            node = "Maize", node.child = "Phenology",
                            node.subchild = "ThermalTime", parm = c("X","Y")) 
```

### File Type (or File Format) Details

APSIM has traditionally generated files that produce simulations using an XML schema. This has been used with APSIM versions of 7.x. With APSIM-X a new format based on JSON files has been implemented and apparently both formats will be supported for the forseeable future. JSON files is less verbose and it is associated with better performance. All distributed APSIM-X examples are 'JSON' now.

The \code{apsimx} package supports inspection and editing of both file types for most components except for the 'Manager'; for which there is partial support. In most instances the parameters that are likely to need inspection and editing are supported but more features will be added in the future as needed. It is possible to test whether a file is XML or JSON.

```{r filetypes}
ex.dir <- auto_detect_apsimx_examples()
apsimx_filetype("Barley", src.dir = ex.dir)
extd.dir <- system.file("extdata", package = "apsimx")
## This is an older XML 'Maize' file which is no longer distributed
apsimx_filetype("Maize_old", src.dir = extd.dir)
```

# In Beta

* inspect_aspimx_replacement
* edit_apsimx_replacement

# Future features

* Create custom function for editing the depth of the soil profile 'edit_apsimx_depth'
* Improve implementation of read_apsimx_all 
* Add tools for optimization and sensitivity analysis.
