---
title: "Parameter Optimization"
author: "Fernando Miguez"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Parameter Optimization}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(apsimx)
library(ggplot2)
apsimx_options(warn.versions = FALSE)
tmp.dir <- tempdir()
```

# Introduction

Often the simulations from a model like APSIM will not be close enough to the observed data. APSIM (Classic and Next Generation) are very complex models with many moving pieces and unexpected interactions among parameters. If you decide that you need to optimize parameters I suggest answering these questions:

1. Why do you need to perform parameter optimization?

2. Which parameters do you need to optimize?

3. Are the observed measurements appropriate for the parameters you want to optimize?

I like to emphasize here the importance of thinking deeply about whether parameter optimization is needed in a given application. Before you use an optimization routine consider the following:

1. Did you control the quality of the weather data ('met' file)?

2. Are the soil properties appropriate for the location being simulated?

3. Is the management information accurate?

Clearly, if there are any problems, issues or inaccuracies which are derived from these three sources of information, then calibrating or tweaking crop parameters to compensate for those errors will not result in a useful model. In concrete terms, as an example: Are you modifying the radition use efficiency parameter to compensate for incorrect planting date or fertilizer inputs? Running an optimization algorithm is easier than having to think hard about the scientific questions, the mechanisms and the structure of the model.

Given this disclaimer, I provide simple functions which connect APSIM with optimization routines. These are  **optim_apsim** and **optim_apsimx**. These functions work best for situations when the function being optimized is smooth (i.e. small changes in the inputs result in small changes in the output without jumps or discontinuities) and the parameters being optimized are continuous. This is: that both input parameters and outputs can take any real value. However, several of the optimization algorithms, such as the default 'Nelder-Mead' work for non-differentiable functions.

I have been testing them and they work as expected. This means that very often they give you the right answer to the wrong question! For example, the optimized value for a parameter might not be within a physical or biological meaningful range if you are not careful. Let's look at the arguments:

## optim_apsim

* file = .apsim file
* src.dir = source directory. Should be the current directory for both this file and the crop file.
* crop.file = typically something like 'Wheat.xml', 'Maize.xml' or similar. You should have also placed an 'ini' under the crop in in the .apsim file and point to this file.
* parm.paths = character vector with the full path to the parameters to be optimized. This can be obtained by using **inspect_apsim** or **inspect_apsim_xml**. The length of the parameter vector should be equal to the number of parameters being optimized (sometimes less is more). For example:

```{r inspect-apsimx-xml-maize}
extd.dir <- system.file("extdata", package = "apsimx")
rue.pth <- inspect_apsim_xml("Maize75.xml", src.dir = extd.dir, parm = "rue")
ext.pth <- inspect_apsim_xml("Maize75.xml", src.dir = extd.dir, parm = "y_extinct_coef")
## To pass them to optim_apsim, combine them
pp <- c(rue.pth, ext.pth)
```
Generates the paths for the radiation use efficiency and extinction coefficient paths. 

* data = data frame with the observed measurement which will be compared to APSIM output. It is important that the names match. This is the names given in the output of an APSIM simulation should have the same names as the column names in this data frame so that they can be aligned. Only one observation is allowed per Date.

* type = type of optimization. The default is to use **optim**. It is also possible to use the **nloptr** package or **mcmc** via the **BayesianTools** package (as of version 1.959).

* weights = weights to apply when multiple variables are being optimized. See help file for more details.

* index = the default is 'Date' this allows the alignment of APSIM output and observed measurements.

* parm.vector.index = some paths point to a vector of parameters, but sometimes only one element of that vector needs to be optimized. In this case provide the index for which element of the vector needs to be optimized. If some of the parameters need this index while others do not, include a zero for those parameter vectors which are not indexed. 

* ... = additional arguments to be passed to **optim**. For example, you can include lower and upper bounds and change the optimization algorithm (in this case 'L-BFGS-B'). See **optim** help file and documentation. Other options for either **nloptr** or **mcmc**.

## optim_apsimx

This function has a very similar structure to **optim_apsim**. The main difference is that at the moment it is required to supply the initial values for the parameters. It is also required to indicate whether each parameter is present in the 'replacement' component. As expalined before, use **inspect_apsimx** or **inspect_apsimx_replacement**.

# Wheat Example

As a simple example, let's imagine we have collected some data on phenology, LAI and biomass of wheat. The data are already in the package and we just need to load it. These data were actually simulated with known parameter values.

## Observed data

```{r obsWheat}
data(obsWheat)
## See the structure
head(obsWheat)
## Only 10 observations
dim(obsWheat)
## Visualize the data
ggplot(obsWheat, aes(Date, Wheat.AboveGround.Wt)) + 
  geom_line() + 
  ggtitle("Biomass (g/m2)")
  
ggplot(obsWheat, aes(Date, Wheat.Leaf.LAI)) + 
  geom_line() +
  ggtitle("LAI")
  
ggplot(obsWheat, aes(Date, Wheat.Phenology.Stage)) + 
  geom_line() +
  ggtitle("Phenology Stages")
```

Notice that the names in the *obsWheat* data frame are identical to the names from the APSIM-X output and this is important later in the optimization so that we can match the variables.

## Before optimization

```{r wheat-sim-b4-opt, echo = FALSE}
sim0 <- read.csv(file.path(extd.dir, "wheat-sim-b4-opt.csv"))
sim0$Date <- as.Date(sim0$Date)
```

```{r sim0-wheat-sim, eval = FALSE}
sim0 <- apsimx("Wheat-opt-ex.apsimx", src.dir = extd.dir, value = "report")
```

```{r sim0-viz}
## Select 
sim0.s <- subset(sim0, 
                 Date > as.Date("2016-09-30") & Date < as.Date("2017-07-01"))
## Visualize before optimization
## phenology
ggplot() + 
  geom_point(data = obsWheat, aes(x = Date, y = Wheat.Phenology.Stage)) +
  geom_line(data = sim0.s, aes(x = Date, y = Wheat.Phenology.Stage)) + 
  ggtitle("Phenology")
## LAI
ggplot() + 
  geom_point(data = obsWheat, aes(x = Date, y = Wheat.Leaf.LAI)) +
  geom_line(data = sim0.s, aes(x = Date, y = Wheat.Leaf.LAI)) + 
  ggtitle("LAI")
## Biomass
ggplot() + 
  geom_point(data = obsWheat, aes(x = Date, y = Wheat.AboveGround.Wt)) +
  geom_line(data = sim0.s, aes(x = Date, y = Wheat.AboveGround.Wt)) + 
    ggtitle("Biomass (g/m2)")
```

## Setting up the optimization

We have observed data and we have ran the model. We notice that the agreement between observed and simulated could be better. For this example we will only be optimizing two parameters: RUE and phyllochron. We could use the function **inspect_apsimx_replacement** to find them. Then we construct the paths.

```{r inspect}
## Finding RUE
inspect_apsimx_replacement("Wheat-opt-ex.apsimx", src.dir = extd.dir,
                           node = "Wheat", 
                           node.child = "Leaf",
                           node.subchild = "Photosynthesis",
                           node.subsubchild = "RUE", 
                           parm = "FixedValue",
                           verbose = FALSE)
## Finding BasePhyllochron
inspect_apsimx_replacement("Wheat-opt-ex.apsimx", src.dir = extd.dir,
                           node = "Wheat", 
                           node.child = "Cultivars",
                           node.subchild = "USA",
                           node.subsubchild = "Yecora", 
                           verbose = FALSE)
## Constructing the paths is straight-forward
pp1 <- "Wheat.Leaf.Photosynthesis.RUE.FixedValue"
pp2 <- "Wheat.Cultivars.USA.Yecora.BasePhyllochron"
```

## Running the optimization

Once we have set up these pieces we can run the optimization. Here, the first argument is the name of the file, the second the directory where it is (*src.dir*), then the paths indicating the parameters to optimize, the observed data (*data*), the weighting method (here = *mean*), whether the parameters are in the replacement part of the simulation and the initial values.

```{r optim-apsimx, eval = FALSE}
## wop is for wheat optimization
wop <- optim_apsimx("Wheat-opt-ex.apsimx", 
                    src.dir = extd.dir, 
                    parm.paths = c(pp1, pp2),
                    data = obsWheat, 
                    weights = "mean",
                    replacement = c(TRUE, TRUE),
                    initial.values = c(1.2, 120))
```

This optimization ran for about 7 minutes (on my laptop). This is the result:

```{r load-wop, echo = FALSE}
data("wheat-opt-ex", package = "apsimx")
```

```{r wop-result}
wop
```

We can see that the optimized values are very close to the known values which were used to simulate the data. The known value for RUE was 1.5 $g \; MJ^{-1}$ and for BasePhyllochron it was 90 GDD.

In addition, it is good practice to compute the Hessian matrix, since it provides information about standard errors and confidence intervals.

```{r optim-apsimx-hessian, eval = FALSE}
## wop is for wheat optimization
wop.h <- optim_apsimx("Wheat-opt-ex.apsimx", 
                      src.dir = extd.dir, 
                      parm.paths = c(pp1, pp2),
                      data = obsWheat, 
                      weights = "mean",
                      replacement = c(TRUE, TRUE),
                      initial.values = c(1.2, 120),
                      hessian = TRUE)
```

This ran for about 8 minutes and it allows for calculation of confidence intervals and standard errors.

```{r wop-result-hessian}
wop.h
```

After optimization there is good agreement between observed and simulated.

```{r wop-result-opt, eval = FALSE}
## We re-run the model because the Wheat-opt-ex.apsimx file 
## is already edited
sim.opt <- apsimx("Wheat-opt-ex.apsimx", src.dir = extd.dir, value = "report")  
sim.opt.s <- subset(sim.opt, 
                    Date > as.Date("2016-09-30") & Date < as.Date("2017-07-01"))
```

```{r import-wop-result, echo = FALSE}
sim.opt.s <- read.csv(file.path(extd.dir, "wheat-sim-opt.csv"))
sim.opt.s$Date <- as.Date(sim.opt.s$Date)
```

```{r wop-result-opt-visualize}
  ## phenology
  ggplot() + 
    geom_point(data = obsWheat, aes(x = Date, y = Wheat.Phenology.Stage)) +
    geom_line(data = sim.opt.s, aes(x = Date, y = Wheat.Phenology.Stage)) + 
    ggtitle("Phenology")
  ## LAI
  ggplot() + 
    geom_point(data = obsWheat, aes(x = Date, y = Wheat.Leaf.LAI)) +
    geom_line(data = sim.opt.s, aes(x = Date, y = Wheat.Leaf.LAI)) + 
    ggtitle("LAI")
  ## Biomass
  ggplot() + 
    geom_point(data = obsWheat, aes(x = Date, y = Wheat.AboveGround.Wt)) +
    geom_line(data = sim.opt.s, aes(x = Date, y = Wheat.AboveGround.Wt)) + 
    ggtitle("Biomass (g/m2)")
```

### Additional considerations and potential issues

The main additional consideration is that it can be a good idea to set up lower and upper bounds on the parameters in terms of how much higher and lower than the default values can be. In a model such as APSIM usually plus or minus 50 percent is enough of a range and even a lower range can often suffice. These can be included as 0.5 and 1.5. With *optim* it is necessary to use method = "L-BFGS-B". There are also similar options when using **nloptr**. This package has many available algorithms to choose from with more modern implementations than the ones available in **optim**. However, one disadvantage of the **nloptr** function is that it does not seem to be able to compute the Hessian numerically. 

* One issue which I have not fully tracked down is an error which appears which indicates a conflict with disk I/O. Again, I need to check but this may be when trying to read and/or write an apsimx file which is also opened in the APSIM GUI application. It is possible that this issue goes away when the APSIM GUI is closed.

* I'm not convinced that using the Hessian (in the way I'm doing it) produces accurate standard errors. It seems to work just fine when there is one variable and perhaps when there is no weighting applied, but it seems that it breaks down with multiple variables and weighting. The **mcmc** method should not have this problem, but it takes a long time to run.
